---
title: "NASA exp3"
author: "Alex Seo"
date: "Mar 15, 2021"
output: pdf_document
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhf{}
- \rhead{Seo}
- \lhead{NASA}
- \cfoot{\thepage}
---

<!-- R chunk setup -->
```{r setup, include=FALSE}
# global chunk options
knitr::opts_chunk$set(echo = TRUE, comment = NA)

# load and attach add-on packages here
library(knitr)
library(qcc)
library(tidyverse)
library(leaps)
library(MVA)
library(MASS)
library(ggfortify)
library(Rtsne)
```

<!-- Page Break -->
\newpage

##Study Description of GLDS-120: 

Experimentation on the International Space Station has reached the stage where repeated and nuanced transcriptome studies are beginning to illuminate the **structural and metabolic differences between plants grown in space compared to plants on the Earth.** 

Genes that are important in setting up the spaceflight responses are being identified; their role in spaceflight physiological adaptation are increasingly understood, and the fact that different genotypes adapt differently is recognized. However, the basic question of whether these spaceflight responses are required for survival has yet to be posed, and the fundamental notion that spaceflight responses may be non-adaptive has yet to be explored.
Therefore the experiments presented here were designed to ask **if portions of the plant spaceflight response can be genetically removed without causing loss of spaceflight survival and without causing increased stress responses.**

The CARA experiment compared the spaceflight transcriptome responses of two Arabidopsis ecotypes, Col-0 and WS, as well as that of a PhyD mutant of Col-0. When grown with the ambient light of the ISS, phyD displayed a significantly reduced spaceflight transcriptome response compared to Col-0, **suggesting that altering the activity of a single gene can actually improve spaceflight adaptation by reducing the transcriptome cost of physiological adaptation.** The WS genotype showed an even simpler spaceflight transcriptome response in the ambient light of the ISS, more **broadly indicating that the plant genotype can be manipulated to reduce the transcriptome cost of plant physiological adaptation to spaceflight and suggesting that genetic manipulation might further reduce, or perhaps eliminate the metabolic cost of spaceflight adaptation.** 

When plants were germinated and then left in the dark on the ISS, the WS genotype actually mounted a larger transcriptome response than Col-0, suggesting that the in-space light environment affects physiological adaptation, which further **implies that manipulating the local habitat can also substantially impact the metabolic cost of spaceflight adaptation**.


```{r}
setwd("C:/Users/LG/Documents/NASA_GeneLab/data")
set.seed(1892)
nasa<-read.csv("GLDS-120_rna_seq_Normalized_Counts.csv")
#CARA_FC<-read.csv("logfc_GLDS120.csv")
##head(nasa)
##summary(nasa)
#Dividing dataset 
#view(nasa)
rna_seq<-nasa[,1]
phyD_FLT<-nasa[,c(1:7)]
phyD_GC<-nasa[,c(1,8:13)]
col_FLT<-nasa[,c(1,14:19)]
col_GC<-nasa[,c(1,20:25)]
WS_FLT<-nasa[,c(1,26:31)]
WS_GC<-nasa[,c(1,32:37)]
#Species
phyD<-cbind(phyD_FLT,phyD_GC[,-1])
col<-cbind(col_FLT,col_GC[,-1])
WS<-cbind(WS_FLT,WS_GC[,-1])

```



```{r}
#col
colnames(col_FLT)<-c('LOCI','FLT_L(1)','FLT_L(2)','FLT_L(3)','FLT_D(1)','FLT_D(2)','FLT_D(3)')
col_FLT<-col_FLT[-c(861, 12958, 23875),]

set.seed(1892)
#PCA
col_FLT.pca<-princomp(col_FLT[,-1])
summary(col_FLT.pca)
a1<-col_FLT.pca$scores[,1]
a2<-col_FLT.pca$scores[,2]
a<-cbind(a1,a2)

pcaplot <- data.frame(PC1 = a1, PC2 = a2)

ggplot(pcaplot, aes(x = PC1, y = PC2)) +
  geom_point()

eigvals<-col_FLT.pca$sdev^2
k<-length(eigvals)
plot(1:k,eigvals,type="b",xlab="i",ylab=expression(lambda[i]))

##PCAbiplot
ggplot2::autoplot(col_FLT.pca,label = F, loadings.label = T)
#which(a1 > 0.75e+05)


```

```{r}
set.seed(1892)
#tsne
col_FLT.tsne<-Rtsne(col_FLT[,-1], dims = 2, perplexity=30, verbose=TRUE, max_iter = 500, check_duplicates = FALSE)
t1<-col_FLT.tsne$Y[,1];t2<-col_FLT.tsne$Y[,2]
t<-cbind(t1,t2)
tsneplot <- data.frame(tsne1 = t1, tsne2 = t2)

ggplot(tsneplot, aes(x = tsne1, y = tsne2)) +
  geom_point()
```


```{r}
##kmeans
n<-dim(a)[1]; k<-6
wss<- rep(0,k); xm<-apply(a,2,mean)
for(i in 1:n){
  wss[1]<- wss[1]+sum((a[i,]-xm)^2)
}
for(i in 2:k){
  model<-kmeans(a,i)
  wss[i]<-sum(model$withinss)
}
plot(1:k,wss,type="b",xlab="Number of clusters", ylab="Within cluster sum of squares",main="Screeplot")

k<-4
set.seed(1892)
km4<-kmeans(a,k)

df <- data.frame(PC1 = a1, PC2 = a2, cluster = factor(km4$cluster))

ggplot(df, aes(x = PC1, y = PC2, color = cluster)) +
  geom_point()


length(col_FLT[which(km4$cluster==1),1]) #43

length(col_FLT[which(km4$cluster==2),1]) #227

length(col_FLT[which(km4$cluster==3),1]) #22398

length(col_FLT[which(km4$cluster==4),1]) #1315
```


```{r}
col_FLT_c1<-col_FLT[which(km4$cluster==2),]
set.seed(1892)
col_FLT.tsne2<-Rtsne(col_FLT_c1[,-1], dims = 2, perplexity=30, verbose=TRUE, max_iter = 500, check_duplicates = FALSE)
t1<-col_FLT.tsne2$Y[,1];t2<-col_FLT.tsne2$Y[,2]
t<-cbind(t1,t2)
tsneplot <- data.frame(tsne1 = t1, tsne2 = t2)

ggplot(tsneplot, aes(x = tsne1, y = tsne2)) +
  geom_point()
```


```{r}
library(mclust)
set.seed(1892)
mclus_default<-Mclust(t)

mclass<-mclus_default$classification
k<-mclus_default$G
(mclus_default$modelName)

df <- data.frame(tsne1 = t1, tsne2 = t2, cluster = factor(mclass))

ggplot(df, aes(x = tsne1, y = tsne2, color = cluster)) +
  geom_point()
```

```{r}
length(col_FLT_c1[which(mclass==1),1]) #3731
col_FLT_tsnem1<-data.frame(col_FLT_c1[which(mclass==1),1])
write.csv(col_FLT_tsnem1, file = "col_FLT_tsnem1.csv")

length(col_FLT_c1[which(mclass==2),1]) #3839
col_FLT_tsnem2<-data.frame(col_FLT_c1[which(mclass==2),1])
write.csv(col_FLT_tsnem2, file = "col_FLT_tsnem2.csv")

length(col_FLT_c1[which(mclass==3),1]) #1436
col_FLT_tsnem3<-data.frame(col_FLT_c1[which(mclass==3),1])
write.csv(col_FLT_tsnem3, file = "col_FLT_tsnem3.csv")

length(col_FLT_c1[which(mclass==4),1]) #4174
col_FLT_tsnem4<-data.frame(col_FLT_c1[which(mclass==4),1])
write.csv(col_FLT_tsnem4, file = "col_FLT_tsnem4.csv")

length(col_FLT_c1[which(mclass==5),1]) #1732
col_FLT_tsnem5<-data.frame(col_FLT_c1[which(mclass==5),1])
write.csv(col_FLT_tsnem5, file = "col_FLT_tsnem5.csv")

length(col_FLT_c1[which(mclass==6),1]) #2126
col_FLT_tsnem6<-data.frame(col_FLT_c1[which(mclass==6),1])
write.csv(col_FLT_tsnem6, file = "col_FLT_tsnem6.csv")

length(col_FLT_c1[which(mclass==7),1]) #481
col_FLT_tsnem7<-data.frame(col_FLT_c1[which(mclass==7),1])
write.csv(col_FLT_tsnem7, file = "col_FLT_tsnem7.csv")

length(col_FLT_c1[which(mclass==8),1]) #3308
col_FLT_tsnem8<-data.frame(col_FLT_c1[which(mclass==8),1])
write.csv(col_FLT_tsnem8, file = "col_FLT_tsnem8.csv")

length(col_FLT_c1[which(mclass==9),1]) #1571
col_FLT_tsnem9<-data.frame(col_FLT_c1[which(mclass==9),1])
write.csv(col_FLT_tsnem9, file = "col_FLT_tsnem9.csv")
```

```{r}
#dbscan
library(dbscan)

kNNdistplot(t, k =  5)
abline(h = 40, lty = 2)

db<-dbscan(t,eps = .5,minPts = 50)
db


df <- data.frame(tsne1 = t1, tsne2 = t2, cluster = factor(db$cluster))


ggplot(df, aes(x = tsne1, y = tsne2, color = cluster)) +
  geom_point()

#length(col_FLT_c1[which(db$cluster==1),1])
#col_FLT_dbscan1<-data.frame(col_FLT_c1[which(db$cluster==1),1])

#write.csv(col_FLT_dbscan1, file = "col_FLT_dbscan1.csv")
```

##GC

```{r}
#col
colnames(col_GC)<-c('LOCI','GC_L(1)','GC_L(2)','GC_L(3)','GC_D(1)','GC_D(2)','GC_D(3)')
col_GC<-col_GC[-c(861, 12958, 23875),]
#12958,23875:GC_D(2), 861: GC_L(2,3)
set.seed(1892)
#PCA
col_GC.pca<-princomp(col_GC[,-1])
summary(col_GC.pca)
a1<-col_GC.pca$scores[,1]
a2<-col_GC.pca$scores[,2]
a<-cbind(a1,a2)

pcaplot <- data.frame(PC1 = a1, PC2 = a2)

ggplot(pcaplot, aes(x = PC1, y = PC2)) +
  geom_point()

eigvals<-col_GC.pca$sdev^2
k<-length(eigvals)
plot(1:k,eigvals,type="b",xlab="i",ylab=expression(lambda[i]))

##PCAbiplot
ggplot2::autoplot(col_GC.pca,label = F, loadings.label = T)#
#which(a1 > 1e+05)

```
```{r}
#tsne
set.seed(1892)
col_GC.tsne<-Rtsne(col_GC[,-1], dims = 2, perplexity=30, verbose=TRUE, max_iter = 500, check_duplicates = FALSE)
t1<-col_GC.tsne$Y[,1];t2<-col_GC.tsne$Y[,2]
t<-cbind(t1,t2)
tsneplot <- data.frame(tsne1 = t1, tsne2 = t2)

ggplot(tsneplot, aes(x = tsne1, y = tsne2)) +
  geom_point()
```

```{r}
##kmeans
n<-dim(a)[1]; k<-6
wss<- rep(0,k); xm<-apply(a,2,mean)
for(i in 1:n){
  wss[1]<- wss[1]+sum((a[i,]-xm)^2)
}
for(i in 2:k){
  model<-kmeans(a,i)
  wss[i]<-sum(model$withinss)
}
plot(1:k,wss,type="b",xlab="Number of clusters", ylab="Within cluster sum of squares",main="Screeplot")

k<-4
set.seed(1892)
km4<-kmeans(a,k)

df <- data.frame(PC1 = a1, PC2 = a2, cluster = factor(km4$cluster))

ggplot(df, aes(x = PC1, y = PC2, color = cluster)) +
  geom_point()


length(col_GC[which(km4$cluster==1),1])

length(col_GC[which(km4$cluster==2),1])

length(col_GC[which(km4$cluster==3),1])

length(col_GC[which(km4$cluster==4),1])
```



```{r}
col_GC_c1<-col_GC[which(km4$cluster==2),]
set.seed(1892)
col_GC.tsne2<-Rtsne(col_GC_c1[,-1], dims = 2, perplexity=30, verbose=TRUE, max_iter = 500, check_duplicates = FALSE)
t1<-col_GC.tsne2$Y[,1];t2<-col_GC.tsne2$Y[,2]
t<-cbind(t1,t2)
tsneplot <- data.frame(tsne1 = t1, tsne2 = t2)

ggplot(tsneplot, aes(x = tsne1, y = tsne2)) +
  geom_point()
```


```{r}
##EM clustering
library(mclust)
set.seed(1892)
mclus_default<-Mclust(t)

mclass<-mclus_default$classification
k<-mclus_default$G
(mclus_default$modelName)

df <- data.frame(tsne1 = t1, tsne2 = t2, cluster = factor(mclass))

ggplot(df, aes(x = tsne1, y = tsne2, color = cluster)) +
  geom_point()
```

```{r}
length(col_GC_c1[which(mclass==1),1]) #2990
col_GC_mclass1<-data.frame(col_GC_c1[which(mclass==1),1])
write.csv(col_GC_mclass1, file = "col_GC_tsnem1.csv")

length(col_GC_c1[which(mclass==2),1]) #2843
col_GC_mclass2<-data.frame(col_GC_c1[which(mclass==2),1])
write.csv(col_GC_mclass2, file = "col_GC_tsnem2.csv")

length(col_GC_c1[which(mclass==3),1]) #1688
col_GC_mclass3<-data.frame(col_GC_c1[which(mclass==3),1])
write.csv(col_GC_mclass3, file = "col_GC_tsnem3.csv")

length(col_GC_c1[which(mclass==4),1]) #2524
col_GC_mclass4<-data.frame(col_GC_c1[which(mclass==4),1])
write.csv(col_GC_mclass4, file = "col_GC_tsnem4.csv")

length(col_GC_c1[which(mclass==5),1]) #2096
col_GC_mclass5<-data.frame(col_GC_c1[which(mclass==5),1])
write.csv(col_GC_mclass5, file = "col_GC_tsnem5.csv")

length(col_GC_c1[which(mclass==6),1]) #480
col_GC_mclass6<-data.frame(col_GC_c1[which(mclass==6),1])
write.csv(col_GC_mclass6, file = "col_GC_tsnem6.csv")

length(col_GC_c1[which(mclass==7),1]) #3984
col_GC_mclass7<-data.frame(col_GC_c1[which(mclass==7),1])
write.csv(col_GC_mclass7, file = "col_GC_tsnem7.csv")

length(col_GC_c1[which(mclass==8),1]) #4009
col_GC_mclass8<-data.frame(col_GC_c1[which(mclass==8),1])
write.csv(col_GC_mclass8, file = "col_GC_tsnem8.csv")

length(col_GC_c1[which(mclass==9),1]) #2296
col_GC_mclass9<-data.frame(col_GC_c1[which(mclass==9),1])
write.csv(col_GC_mclass9, file = "col_GC_tsnem9.csv")
```


```{r}
#dbscan
library(dbscan)

kNNdistplot(t, k =  5)
abline(h = 50, lty = 2)

db<-dbscan(t,eps = .5,minPts = 50)
db


df <- data.frame(tsne1 = t1, tsne2 = t2, cluster = factor(db$cluster))


ggplot(df, aes(x = tsne1, y = tsne2, color = cluster)) +
  geom_point()

#length(col_GC_c1[which(db$cluster==1),1])
#col_GC_dbscan1<-data.frame(col_GC_c1[which(db$cluster==1),1])

#write.csv(col_GC_dbscan1, file = "col_GC_dbscan1.csv")
```

