---
title: "NASA exp3"
author: "Alex Seo"
date: "Mar 15, 2021"
output: pdf_document
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhf{}
- \rhead{Seo}
- \lhead{NASA}
- \cfoot{\thepage}
---

<!-- R chunk setup -->
```{r setup, include=FALSE}
# global chunk options
knitr::opts_chunk$set(echo = TRUE, comment = NA)

# load and attach add-on packages here
library(knitr)
library(qcc)
library(tidyverse)
library(leaps)
library(MVA)
library(MASS)
library(ggfortify)
library(Rtsne)
```

<!-- Page Break -->
\newpage

##Study Description of GLDS-120: 

Experimentation on the International Space Station has reached the stage where repeated and nuanced transcriptome studies are beginning to illuminate the **structural and metabolic differences between plants grown in space compared to plants on the Earth.** 

Genes that are important in setting up the spaceflight responses are being identified; their role in spaceflight physiological adaptation are increasingly understood, and the fact that different genotypes adapt differently is recognized. However, the basic question of whether these spaceflight responses are required for survival has yet to be posed, and the fundamental notion that spaceflight responses may be non-adaptive has yet to be explored.
Therefore the experiments presented here were designed to ask **if portions of the plant spaceflight response can be genetically removed without causing loss of spaceflight survival and without causing increased stress responses.**

The CARA experiment compared the spaceflight transcriptome responses of two Arabidopsis ecotypes, Col-0 and WS, as well as that of a PhyD mutant of Col-0. When grown with the ambient light of the ISS, phyD displayed a significantly reduced spaceflight transcriptome response compared to Col-0, **suggesting that altering the activity of a single gene can actually improve spaceflight adaptation by reducing the transcriptome cost of physiological adaptation.** The WS genotype showed an even simpler spaceflight transcriptome response in the ambient light of the ISS, more **broadly indicating that the plant genotype can be manipulated to reduce the transcriptome cost of plant physiological adaptation to spaceflight and suggesting that genetic manipulation might further reduce, or perhaps eliminate the metabolic cost of spaceflight adaptation.** 

When plants were germinated and then left in the dark on the ISS, the WS genotype actually mounted a larger transcriptome response than Col-0, suggesting that the in-space light environment affects physiological adaptation, which further **implies that manipulating the local habitat can also substantially impact the metabolic cost of spaceflight adaptation**.


```{r}
setwd("C:/Users/LG/Documents/NASA_GeneLab/data")
set.seed(1892)
nasa<-read.csv("GLDS-120_rna_seq_Normalized_Counts.csv")

##head(nasa)
##summary(nasa)
#Dividing dataset 
#view(nasa)
rna_seq<-nasa[,1]
phyD_FLT<-nasa[,c(1:7)]
phyD_GC<-nasa[,c(1,8:13)]
col_FLT<-nasa[,c(1,14:19)]
col_GC<-nasa[,c(1,20:25)]
WS_FLT<-nasa[,c(1,26:31)]
WS_GC<-nasa[,c(1,32:37)]
#Species
phyD<-cbind(phyD_FLT,phyD_GC[,-1])
col<-cbind(col_FLT,col_GC[,-1])
WS<-cbind(WS_FLT,WS_GC[,-1])

```

```{r}
#phyD
colnames(phyD_FLT)<-c('LOCI','FLT_L(1)','FLT_L(2)','FLT_L(3)','FLT_D(1)','FLT_D(2)','FLT_D(3)')
phyD_FLT<-phyD_FLT[-c(861, 12958, 23875),]

set.seed(1892)
#PCA
phyD_FLT.pca<-princomp(phyD_FLT[,-1])
summary(phyD_FLT.pca)
a1<-phyD_FLT.pca$scores[,1]
a2<-phyD_FLT.pca$scores[,2]
a<-cbind(a1,a2)

pcaplot <- data.frame(PC1 = a1, PC2 = a2)

ggplot(pcaplot, aes(x = PC1, y = PC2)) +
  geom_point()

eigvals<-phyD_FLT.pca$sdev^2
k<-length(eigvals)
plot(1:k,eigvals,type="b",xlab="i",ylab=expression(lambda[i]))

##PCAbiplot
ggplot2::autoplot(phyD_FLT.pca,label = F, loadings.label = T)
#which(a1 > 1e+05)

```


```{r}
set.seed(1892)
#tsne
phyD_FLT.tsne<-Rtsne(phyD_FLT[,-1], dims = 2, perplexity=30, verbose=TRUE, max_iter = 500, check_duplicates = FALSE)
t1<-phyD_FLT.tsne$Y[,1];t2<-phyD_FLT.tsne$Y[,2]
t<-cbind(t1,t2)
tsneplot <- data.frame(tsne1 = t1, tsne2 = t2)

ggplot(tsneplot, aes(x = tsne1, y = tsne2)) +
  geom_point()
```

```{r}
##kmeans
n<-dim(a)[1]; k<-6
wss<- rep(0,k); xm<-apply(a,2,mean)
for(i in 1:n){
  wss[1]<- wss[1]+sum((a[i,]-xm)^2)
}
for(i in 2:k){
  model<-kmeans(a,i)
  wss[i]<-sum(model$withinss)
}
plot(1:k,wss,type="b",xlab="Number of clusters", ylab="Within cluster sum of squares",main="Screeplot")

k<-4
set.seed(1892)
km4<-kmeans(a,k)

df <- data.frame(PC1 = a1, PC2 = a2, cluster = factor(km4$cluster))

ggplot(df, aes(x = PC1, y = PC2, color = cluster)) +
  geom_point()


length(phyD_FLT[which(km4$cluster==1),1])

length(phyD_FLT[which(km4$cluster==2),1])

length(phyD_FLT[which(km4$cluster==3),1])

length(phyD_FLT[which(km4$cluster==4),1])
```



```{r}
phyD_FLT_c1<-phyD_FLT[which(km4$cluster==2),]
set.seed(1892)
phyD_FLT.tsne2<-Rtsne(phyD_FLT_c1[,-1], dims = 2, perplexity=30, verbose=TRUE, max_iter = 500, check_duplicates = FALSE)
t1<-phyD_FLT.tsne2$Y[,1];t2<-phyD_FLT.tsne2$Y[,2]
t<-cbind(t1,t2)
tsneplot <- data.frame(tsne1 = t1, tsne2 = t2)

ggplot(tsneplot, aes(x = tsne1, y = tsne2)) +
  geom_point()
```


```{r}
library(mclust)
set.seed(1892)
mclus_default<-Mclust(t)

mclass<-mclus_default$classification
k<-mclus_default$G
(mclus_default$modelName)

df <- data.frame(tsne1 = t1, tsne2 = t2, cluster = factor(mclass))

ggplot(df, aes(x = tsne1, y = tsne2, color = cluster)) +
  geom_point()
```

```{r}
length(phyD_FLT_c1[which(mclass==1),1])
phyD_FLT_mclass1<-data.frame(phyD_FLT_c1[which(mclass==1),1])#2864
write.csv(phyD_FLT_mclass1, file = "phyD_FLT_tsnem1.csv")

length(phyD_FLT_c1[which(mclass==2),1])
phyD_FLT_mclass2<-data.frame(phyD_FLT_c1[which(mclass==2),1])#3392
write.csv(phyD_FLT_mclass2, file = "phyD_FLT_tsnem2.csv")

length(phyD_FLT_c1[which(mclass==3),1])
phyD_FLT_mclass3<-data.frame(phyD_FLT_c1[which(mclass==3),1])#2873
write.csv(phyD_FLT_mclass3, file = "phyD_FLT_tsnem3.csv")

length(phyD_FLT_c1[which(mclass==4),1])
phyD_FLT_mclass4<-data.frame(phyD_FLT_c1[which(mclass==4),1])#1652
write.csv(phyD_FLT_mclass4, file = "phyD_FLT_tsnem4.csv")

length(phyD_FLT_c1[which(mclass==5),1])
phyD_FLT_mclass5<-data.frame(phyD_FLT_c1[which(mclass==5),1])#590
write.csv(phyD_FLT_mclass5, file = "phyD_FLT_tsnem5.csv")

length(phyD_FLT_c1[which(mclass==6),1])
phyD_FLT_mclass6<-data.frame(phyD_FLT_c1[which(mclass==6),1])#2585
write.csv(phyD_FLT_mclass6, file = "phyD_FLT_tsnem6.csv")

length(phyD_FLT_c1[which(mclass==7),1])
phyD_FLT_mclass7<-data.frame(phyD_FLT_c1[which(mclass==7),1])#2950
write.csv(phyD_FLT_mclass7, file = "phyD_FLT_tsnem7.csv")

length(phyD_FLT_c1[which(mclass==8),1])
phyD_FLT_mclass8<-data.frame(phyD_FLT_c1[which(mclass==8),1])#4158
write.csv(phyD_FLT_mclass8, file = "phyD_FLT_tsnem8.csv")

length(phyD_FLT_c1[which(mclass==9),1])
phyD_FLT_mclass9<-data.frame(phyD_FLT_c1[which(mclass==9),1])#1248
write.csv(phyD_FLT_mclass9, file = "phyD_FLT_tsnem9.csv")
```

```{r}
#dbscan
library(dbscan)

kNNdistplot(t, k =  5)
abline(h = 40, lty = 2)

db<-dbscan(t,eps = .5,minPts = 50)
db


df <- data.frame(tsne1 = t1, tsne2 = t2, cluster = factor(db$cluster))


ggplot(df, aes(x = tsne1, y = tsne2, color = cluster)) +
  geom_point()

```

##GC

```{r}
#phyD
colnames(phyD_GC)<-c('LOCI','GC_L(1)','GC_L(2)','GC_L(3)','GC_D(1)','GC_D(2)','GC_D(3)')
phyD_GC<-phyD_GC[-c(861, 12958, 23875),]
set.seed(1892)
#PCA
phyD_GC.pca<-princomp(phyD_GC[,-1])
summary(phyD_GC.pca)
a1<-phyD_GC.pca$scores[,1]
a2<-phyD_GC.pca$scores[,2]
a<-cbind(a1,a2)

pcaplot <- data.frame(PC1 = a1, PC2 = a2)

ggplot(pcaplot, aes(x = PC1, y = PC2)) +
  geom_point()

eigvals<-phyD_GC.pca$sdev^2
k<-length(eigvals)
plot(1:k,eigvals,type="b",xlab="i",ylab=expression(lambda[i]))

##PCAbiplot
ggplot2::autoplot(phyD_GC.pca,label = F, loadings.label = T)#
#which(a1 > 1e+05)

```

```{r}
#tsne
set.seed(1892)
phyD_GC.tsne<-Rtsne(phyD_GC[,-1], dims = 2, perplexity=30, verbose=TRUE, max_iter = 500, check_duplicates = FALSE)
t1<-phyD_GC.tsne$Y[,1];t2<-phyD_GC.tsne$Y[,2]
t<-cbind(t1,t2)
tsneplot <- data.frame(tsne1 = t1, tsne2 = t2)

ggplot(tsneplot, aes(x = tsne1, y = tsne2)) +
  geom_point()
```

```{r}
##kmeans
n<-dim(a)[1]; k<-6
wss<- rep(0,k); xm<-apply(a,2,mean)
for(i in 1:n){
  wss[1]<- wss[1]+sum((a[i,]-xm)^2)
}
for(i in 2:k){
  model<-kmeans(a,i)
  wss[i]<-sum(model$withinss)
}
plot(1:k,wss,type="b",xlab="Number of clusters", ylab="Within cluster sum of squares",main="Screeplot")

k<-4
set.seed(1892)
km4<-kmeans(a,k)

df <- data.frame(PC1 = a1, PC2 = a2, cluster = factor(km4$cluster))

ggplot(df, aes(x = PC1, y = PC2, color = cluster)) +
  geom_point()


length(phyD_GC[which(km4$cluster==1),1])

length(phyD_GC[which(km4$cluster==2),1])

length(phyD_GC[which(km4$cluster==3),1])

length(phyD_GC[which(km4$cluster==4),1])
```




```{r}
phyD_GC_c1<-phyD_GC[which(km4$cluster==2),]
set.seed(1892)
phyD_GC.tsne2<-Rtsne(phyD_GC_c1[,-1], dims = 2, perplexity=30, verbose=TRUE, max_iter = 500, check_duplicates = FALSE)
t1<-phyD_GC.tsne2$Y[,1];t2<-phyD_GC.tsne2$Y[,2]
t<-cbind(t1,t2)
tsneplot <- data.frame(tsne1 = t1, tsne2 = t2)

ggplot(tsneplot, aes(x = tsne1, y = tsne2)) +
  geom_point()
```


```{r}
##EM clustering
library(mclust)
set.seed(1892)
mclus_default<-Mclust(t)

mclass<-mclus_default$classification
k<-mclus_default$G
(mclus_default$modelName)

df <- data.frame(tsne1 = t1, tsne2 = t2, cluster = factor(mclass))

ggplot(df, aes(x = tsne1, y = tsne2, color = cluster)) +
  geom_point()
```

```{r}
length(phyD_GC_c1[which(mclass==1),1])
phyD_GC_mclass1<-data.frame(phyD_GC_c1[which(mclass==1),1])#4957
write.csv(phyD_GC_mclass1, file = "phyD_GC_mclass1.csv")

length(phyD_GC_c1[which(mclass==2),1])
phyD_GC_mclass2<-data.frame(phyD_GC_c1[which(mclass==2),1])#3485
write.csv(phyD_GC_mclass2, file = "phyD_GC_mclass2.csv")

length(phyD_GC_c1[which(mclass==3),1])
phyD_GC_mclass3<-data.frame(phyD_GC_c1[which(mclass==3),1])#1570
write.csv(phyD_GC_mclass3, file = "phyD_GC_mclass3.csv")

length(phyD_GC_c1[which(mclass==4),1])
phyD_GC_mclass4<-data.frame(phyD_GC_c1[which(mclass==4),1])#3632
write.csv(phyD_GC_mclass4, file = "phyD_GC_mclass4.csv")

length(phyD_GC_c1[which(mclass==5),1])
phyD_GC_mclass5<-data.frame(phyD_GC_c1[which(mclass==5),1])#1897
write.csv(phyD_GC_mclass5, file = "phyD_GC_mclass5.csv")

length(phyD_GC_c1[which(mclass==6),1])
phyD_GC_mclass6<-data.frame(phyD_GC_c1[which(mclass==6),1])#3052
write.csv(phyD_GC_mclass6, file = "phyD_GC_mclass6.csv")

length(phyD_GC_c1[which(mclass==7),1])
phyD_GC_mclass7<-data.frame(phyD_GC_c1[which(mclass==7),1])#1028
write.csv(phyD_GC_mclass7, file = "phyD_GC_mclass7.csv")

length(phyD_GC_c1[which(mclass==8),1])
phyD_GC_mclass8<-data.frame(phyD_GC_c1[which(mclass==8),1])#2241
write.csv(phyD_GC_mclass8, file = "phyD_GC_mclass8.csv")

length(phyD_GC_c1[which(mclass==9),1])
phyD_GC_mclass9<-data.frame(phyD_GC_c1[which(mclass==9),1])#614
write.csv(phyD_GC_mclass9, file = "phyD_GC_mclass9.csv")
```


```{r}
#dbscan
library(dbscan)

kNNdistplot(t, k =  5)
abline(h = 50, lty = 2)

db<-dbscan(t,eps = .5,minPts = 50)
db


df <- data.frame(tsne1 = t1, tsne2 = t2, cluster = factor(db$cluster))


ggplot(df, aes(x = tsne1, y = tsne2, color = cluster)) +
  geom_point()


```

